```{r setup, include=FALSE,echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tibble.print_max = Inf)
```

```{r,echo=FALSE}
library(tidyverse)
library(glue)
```
```{r}
cost_energy_boxplot_prob <- function(file_path, show_points = TRUE,targetCol="cost",thresh="()",opt="",fallback=FALSE,target) {
  library(readr)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  
  lines <- readLines(file_path, warn = FALSE)
  start_idx <- grep("^#\\s*all_solutions\\s*$", lines)

  if (length(start_idx) == 0) {
    stop("no # all_solution find")
  }

  all_lines <- lines[(start_idx + 1):length(lines)]
  all_lines <- all_lines[all_lines != ""]

  df <- read_csv(
    I(paste(all_lines, collapse = "\n")),
    show_col_types = FALSE
  )

  
 ## df <- read_csv(file_path, show_col_types = FALSE, text = paste(all_lines, collapse = "\n"),
## stringsAsFactors = FALSE)

  if (!("probability" %in% names(df)) && ("prob" %in% names(df))) {
    df <- df %>% rename(probability = prob)
  }

  df <- df %>%
    transmute(
      used_fallback=as.logical(used_fallback),
      cost = as.numeric(.data[[targetCol]]),
      energy = as.numeric(energy),
      probability = as.numeric(probability)
    ) %>%
    filter(is.finite(energy), probability > 0,used_fallback == fallback)

  w0 <- min(df$probability)
  df_exp <- df %>%
    mutate(w = pmax(1L, round(probability / w0))) %>%
    uncount(w)

  p <- ggplot(df_exp, aes(x = factor(cost), y = energy)) +
    geom_boxplot() +
    theme_bw() +
    labs(x = targetCol, y = "Energy",
         title = glue("Energy vs {targetCol}; thres: {thresh};opt:{opt}; used_fallback={fallback} "))

  if (show_points) {
    br <- sort(unique(df$probability))
    p <- p +
      geom_point(
        data = df,
        aes(size = probability),
        alpha = 0.5,
        position = position_jitter(width = 0.15, height = 0)
      ) +
      scale_size_continuous(breaks = br, labels = br, name = "probability")
  }
  cost_counts <- df %>%
  count(cost, name = "n_rows") %>%
  mutate(pct = n_rows / sum(n_rows)) %>%
  arrange(cost)
  print(glue("Matrix for used_fallback={fallback}"))
  print(cost_counts)
  print(p)

return(list(
  cost_counts[cost_counts$cost == target, ]$pct
))  }

```
```{r}
# function for plotting energies from two csvs (three with file_multi) in the
# same graph along with a horizontal theoretical-energy reference line.
# file_multi: winning run's energy_per_iteration_*.csv from an
# IBMQExperiments.py --spiq_starts K run (the unsuffixed file, not _start{j}).
# random_offset: added to the random-initialization energies; use the Ising
# offset for runs logged on the Ising scale (before the QUBO-scale fix).
# max_iter: only iterations <= max_iter are plotted (and used for the y range),
# cutting the flat post-convergence tail; NULL plots everything.
# multi_label: legend text for the file_multi curve.

plot_energy_comparison <- function(
    file_random,
    file_spiq,
    reference_energy,
    file_multi = NULL,
    max_iter = 1000,
    random_offset = 0,
    multi_label = "seeded multi-start spiq initialization"
) {
  
  # Read CSVs
  df_random <- read.csv(file_random)
  df_spiq   <- read.csv(file_spiq)
  df_multi  <- if (!is.null(file_multi)) read.csv(file_multi) else NULL
  df_random$energy <- df_random$energy + random_offset
  
  # Sanity checks
  required_cols <- c("iteration", "energy")
  
  if (!all(required_cols %in% colnames(df_random))) {
    stop("Random file must contain 'iteration' and 'energy' columns")
  }
  
  if (!all(required_cols %in% colnames(df_spiq))) {
    stop("SPIQ file must contain 'iteration' and 'energy' columns")
  }
  
  if (!is.null(df_multi) && !all(required_cols %in% colnames(df_multi))) {
    stop("Multi-start file must contain 'iteration' and 'energy' columns")
  }
  
  # Zoom into the first max_iter iterations
  if (!is.null(max_iter)) {
    df_random <- df_random[df_random$iteration <= max_iter, ]
    df_spiq   <- df_spiq[df_spiq$iteration <= max_iter, ]
    if (!is.null(df_multi)) df_multi <- df_multi[df_multi$iteration <= max_iter, ]
  }
  
  # Combined x range
  x_range <- range(
    c(df_random$iteration, df_spiq$iteration, df_multi$iteration),
    na.rm = TRUE
  )
  
  # Combined y range INCLUDING reference line
  y_range <- range(
    c(
      df_random$energy,
      df_spiq$energy,
      df_multi$energy,
      reference_energy
    ),
    na.rm = TRUE
  )
  
  # Padding
  y_pad <- 0.05 * diff(y_range)
  
  if (y_pad == 0) {
    y_pad <- 1
  }
  
  y_range <- c(
    y_range[1] - y_pad,
    y_range[2] + y_pad
  )
  
  # Two stacked panels: the plot on top, the legend in its own strip below,
  # so the legend never overlaps the plot at any figure size
  old_par <- par(no.readonly = TRUE)
  on.exit(par(old_par))
  layout(matrix(1:2, ncol = 1), heights = c(5, 1))
  par(mar = c(4, 4, 4, 2) + 0.1)
  
  # Main plot
  plot(
    df_random$iteration,
    df_random$energy,
    type = "l",
    col = "blue",
    lwd = 0.5,
    xlim = x_range,
    ylim = y_range,
    xlab = "Iteration",
    ylab = "Energy",
    main = "Energy vs Iteration (QAOA)"
  )
  
  # SPIQ curve
  lines(
    df_spiq$iteration,
    df_spiq$energy,
    col = "red",
    lwd = 0.5
  )
  
  # Multi-start SPIQ curve
  if (!is.null(df_multi)) {
    lines(
      df_multi$iteration,
      df_multi$energy,
      col = "purple",
      lwd = 0.5
    )
    
    points(
      df_multi$iteration,
      df_multi$energy,
      col = "purple",
      pch = 16,
      cex = 0.15
    )
  }
  
  # Horizontal reference line
  abline(
    h = reference_energy,
    col = "green3",
    lwd = 1.5,
    lty = 2   # dotted/dashed
  )
  
  # Optional points
  points(
    df_random$iteration,
    df_random$energy,
    col = "blue",
    pch = 16,
    cex = 0.15
  )
  
  points(
    df_spiq$iteration,
    df_spiq$energy,
    col = "red",
    pch = 16,
    cex = 0.15
  )
  
  # Grid
  grid()
  
  # Legend panel (multi-start entry only when that curve is drawn)
  par(mar = c(0, 0, 0, 0))
  plot.new()
  has_multi <- !is.null(df_multi)
  legend(
    "center",
    ncol = 2,
    legend = c(
      "random initialization",
      "spiq initialization",
      if (has_multi) multi_label,
      "ground state energy"
    ),
    col = c("blue", "red", if (has_multi) "purple", "green3"),
    lwd = c(1, 1, if (has_multi) 1, 1.5),
    lty = c(1, 1, if (has_multi) 1, 2),
    pch = c(16, 16, if (has_multi) 16, NA),
    bty = "n"
  )
}
```

Seeded diverse multi-start SPIQ (Week93-95): 8 independently seeded GAs
(seed 0), a pool of every GA's elite plus per-generation bests, deduplicated by
quantum state, energy window 0.25, max-min Hamming selection of 10 candidates,
3 QAOA starts (COBYLA, 1000 iterations, reps 2). The purple curve is the winning
start. The random and single-start SPIQ baselines are the same as in multi-spiq.rmd.
See RESULTS_INDEX.md and CHANGELOG.md.

Problem 1: 3 relations, 2 predicates (5 qubits)
```{r}
# Energy comparison - 3 table problem 1 (seeded multi-start: Week93)
plot_energy_comparison("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/uninitialized/energy/Week71/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       "/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/energy/Week90/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       reference_energy=1.031,
                       file_multi="/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/energy/Week93/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input0/trial1/energy_per_iteration_1000_COBYLA_2_1.csv",
                       max_iter=200)
```

Problem 2: 4 relations, 2 predicates (12 qubits)
```{r}
# Energy comparison - 4 table problem 2 (seeded multi-start: Week94)
plot_energy_comparison("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/uninitialized/energy/Week74/Week84/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       "/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/spiq/energy/Week88/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       reference_energy=4.766,
                       file_multi="/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/energy/Week94/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input1/trial1/energy_per_iteration_1000_COBYLA_2_1.csv",
                       max_iter=400)
```

Problem 3: 4 relations, 6 predicates (20 qubits)
```{r}
# Energy comparison - 4 table problem 3 (seeded multi-start: Week95)
plot_energy_comparison("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/uninitialized/energy/Week24/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       "/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/spiq/energy/Week60/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_10000/reps_2/COBYLA/input0/trial1/energy_per_iteration_10000_COBYLA_2_1.csv",
                       reference_energy=2.725,
                       file_multi="/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/energy/Week95/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input2/trial1/energy_per_iteration_1000_COBYLA_2_1.csv",
                       max_iter=600,
                       random_offset=25.799857164004695)
```


## Optimal vs valid (raw decoding vs fallback) ratios: seeded multi-start

Each sampled bitstring is decoded twice. `used_fallback = FALSE` is the raw join
order; `used_fallback = TRUE` is the order after the fallback repair. For each run,
`cost_energy_boxplot_prob` prints the cost distribution for one decoding and
returns the share of samples at the optimal cost (`target`). The optimal costs were
brute-forced over every left-deep order with
`Scripts/Postprocessing.get_costs_for_leftdeep_tree`: P1 = 15, P2 = 315, P3 = 39.
"Final" is the re-sampled best state (`trial_1/readout_summary.csv`). "First
iteration" is evaluation 1 of the winning start (`readout_summary_eval1.csv`), i.e.
the SPIQ starting point.

```{r}
# Collects the optimal ratio of every boxplot below for the summary table
ratio_rows <- list()
record_ratio <- function(problem, stage, fallback, res, target) {
  pct <- unlist(res)
  ratio_rows[[length(ratio_rows) + 1]] <<- data.frame(
    problem = problem, stage = stage, used_fallback = fallback, optimal_cost = target,
    optimal_ratio = if (length(pct) == 0) 0 else pct[1]
  )
}
```

3 table problem 1 (seeded multi-start SPIQ, Week93) - final
```{r}
# Final QAOA seeded multi-start SPIQ; 3 table - problem1, trial1 (Week93); optimal cost 15

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week93/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=15)
    record_ratio("P1", "final", FALSE, res, 15)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week93/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=15)
    record_ratio("P1", "final", TRUE, res, 15)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

3 table problem 1 (seeded multi-start SPIQ, Week93) - first iteration
```{r}
# First iteration QAOA seeded multi-start SPIQ; 3 table - problem1, trial1 (Week93); optimal cost 15

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week93/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input0/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=15)
    record_ratio("P1", "first iteration", FALSE, res, 15)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week93/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input0/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=15)
    record_ratio("P1", "first iteration", TRUE, res, 15)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

4 table problem 2 (seeded multi-start SPIQ, Week94) - final
```{r}
# Final QAOA seeded multi-start SPIQ; 4 table - problem2, trial1 (Week94); optimal cost 315

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week94/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=315)
    record_ratio("P2", "final", FALSE, res, 315)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week94/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=315)
    record_ratio("P2", "final", TRUE, res, 315)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

4 table problem 2 (seeded multi-start SPIQ, Week94) - first iteration
```{r}
# First iteration QAOA seeded multi-start SPIQ; 4 table - problem2, trial1 (Week94); optimal cost 315

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week94/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input1/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=315)
    record_ratio("P2", "first iteration", FALSE, res, 315)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week94/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input1/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=315)
    record_ratio("P2", "first iteration", TRUE, res, 315)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

4 table problem 3 (seeded multi-start SPIQ, Week95) - final
```{r}
# Final QAOA seeded multi-start SPIQ; 4 table - problem3, trial1 (Week95); optimal cost 39

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week95/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=39)
    record_ratio("P3", "final", FALSE, res, 39)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week95/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/trial_1/readout_summary.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=39)
    record_ratio("P3", "final", TRUE, res, 39)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

4 table problem 3 (seeded multi-start SPIQ, Week95) - first iteration
```{r}
# First iteration QAOA seeded multi-start SPIQ; 4 table - problem3, trial1 (Week95); optimal cost 39

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week95/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input2/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = FALSE, target=39)
    record_ratio("P3", "first iteration", FALSE, res, 39)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (raw) = {round(unlist(res)[1], 4)}"))

res <- cost_energy_boxplot_prob(glue("/Users/divyashekar/Documents/Sem3/QuantumDB/sigmod23-reproduction/outputs/multi-spiq/Week95/ExperimentalAnalysis/IBMQ/QPUPerformance/Results/CPU_Data/iterations_1000/reps_2/COBYLA/input2/trial1/readout_summary_eval1.csv"),targetCol="cost",thresh=150,opt="COBYLA",fallback = TRUE, target=39)
    record_ratio("P3", "first iteration", TRUE, res, 39)
    print(glue("optimizer:COBYLA; input thres:150; costType:cost; optimal ratio (fallback) = {round(unlist(res)[1], 4)}"))
```

Summary: optimal ratio per problem, stage and decoding
```{r}
ratio_summary <- do.call(rbind, ratio_rows)
knitr::kable(ratio_summary, digits = 4,
             caption = "Seeded multi-start SPIQ (Week93-95): share of samples at the optimal join-order cost")
```
